%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Glossary.tex                                        %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Andre C. Marta                                           %
%     Last modified : 30 Oct 2012                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% The definitions can be placed anywhere in the document body
% and their order is sorted by <symbol> automatically when
% calling makeindex in the makefile
%
% The \glossary command has the following syntax:
%
% \glossary{entry}
%
% The \nomenclature command has the following syntax:
%
% \nomenclature[<prefix>]{<symbol>}{<description>}
%
% where <prefix> is used for fine tuning the sort order,
% <symbol> is the symbol to be described, and <description> is
% the actual description.
% ----------------------------------------------------------------------

\glossary{name={\textbf{CRM}},description={Customer Relationship Management is a system for managing a company’s interactions with current and future customers.}}

\glossary{name={\textbf{ANN}},description={Artificial Neural Network}}

\glossary{name={\textbf{SVM}},description={Support Vector Machines are supervised learning models with associated learning algorithms that analyze data and recognize patterns, used for classification and regression analysis.}}

\glossary{name={\textbf{HMM}},description={Hidden Markov Model is a statistical Markov model in which the system being modeled is assumed to be a Markov process with unobserved, in other words, hidden states.}}

\glossary{name={\textbf{MLP}},description={ MultiLayer Perceptron is a feedforward artificial neural network model that maps sets of input data onto a set of appropriate outputs.}}

\glossary{name={\textbf{SEER}},description={Surveillance, Epidemiology, and End Results program.}}

\glossary{name={\textbf{CART}},description={Classification And Regression Tree}}

\glossary{name={\textbf{TTR}},description={Time To Recur}}

\glossary{name={\textbf{ROC}},description={Receiver Operating Characteristic curve is a graphical plot which illustrates the performance of a binary classifier system as its discrimination threshold is varied.}}

\glossary{name={\textbf{MARS}},description={Multivariate Adaptive Regression Splines is a non-parametric regression technique and can be seen as an extension of linear models that automatically models non-linearities and interactions between variables.}}

\glossary{name={\textbf{RSA}},description={Recurrence Surface to Approximate}}

\glossary{name={\textbf{RRLC}},description={Ridge Regression with Linear Constraints}}

\glossary{name={\textbf{AUC}},description={Area Under Curve is a metric calculated with the ROC Curve. It is equal to the probability that a classifier will rank a randomly chosen positive instance higher than a randomly chosen negative one (assuming 'positive' ranks higher than 'negative').}}

\glossary{name={\textbf{PNN}},description={Probabilistic Neural Network is a feedforward neural network, which was derived from the Bayesian network and a statistical algorithm called Kernel Fisher discriminant analysis.}}

\glossary{name={\textbf{k-NN}},description={k-Nearest Neighbor algorithm is a non-parametric method used for classification and regression.}}

\glossary{name={\textbf{ALS}},description={Amyotrophic Lateral Sclerosis.}}



